```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "assets/images/",
  out.width = "100%"
)
```
The core suite of r2dii packages ([r2dii.data](https://2degreesinvesting.github.io/r2dii.data/), [r2dii.match](https://2degreesinvesting.github.io/r2dii.match/) and [r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/)), are now published on CRAN!

The r2dii suite was developed to implement the core functionality of the [Paris Agreement Capital Transition Assessment (PACTA)](https://www.transitionmonitor.com/) in R. PACTA is a methodology which allows financial institutions to aggregate climate-related data associated to their portfolios, and compare the result against climate scenarios.

## Installation

You can easily install all three packages directly from CRAN using:

```{r eval = FALSE}
install.packages(
  c(
    "r2dii.data", 
    "r2dii.match", 
    "r2dii.analysis"
    )
  )
```

To get the latest development version of each of the packages, you can install directly from GitHub using:

``` {r eval = FALSE}
devtools::install_github("2DegreesInvesting/r2dii.data")
```

**Warning:** Development versions are experimental and may be unstable. Proceed with caution.
{: .notice--warning}

## r2dii.data

[r2dii.data](https://2degreesinvesting.github.io/r2dii.data/) provides a series of fake datasets. These are mainly meant to be used as a tool to learn how the methodology works, and also to provide an easy sandbox for testing. The built-in `data_dictionary` provides a view of all datasets in the package, along with variable definitions

```{r}
library(dplyr)
library(r2dii.data)

data_dictionary %>%
  distinct(dataset)
```

Some of the most useful datasets are: 
* `loanbook_demo`: a loanbook dataset which can be used as a template to format real loanbook/ portfolio data. 
* `ald_demo`: a so-called "asset-level" dataset, containing crucial climate data at the asset level. (eg. power-plant capacity data, automotive car production, steel company emission factors)
* `scenario_demo_2020`: a fake climate scenario, offering technology pathways that would likely limit warming to less than 2 degrees celsius. 

## r2dii.match

[r2dii.match](https://2degreesinvesting.github.io/r2dii.match/) provides the tools necessary to match the counter-parties in a loanbook to the climate data of the assets that they own. 

Owing to the fact that many lending portfolios are exposed to mid- and small- size companies, which are often not publicly-listed, matching data can be tricky. To achieve these links, we have written wrappers around some common fuzzy matching algorithms, which play nicely with formatted `loanbook`'s and `ald`'s. 

You can run the matching algorithm on sample data provided by `r2dii.data`:

```{r}
library(r2dii.match)

matched <- match_name(loanbook_demo, ald_demo)

matched %>%
  select(name, sector, name_ald, sector_ald, score) %>%
  head()
```

After the initial matching, you must manually verify which of the matches you would like to keep. To do so, simply save `matched` as a .csv, and open in excel or similar. If you would like to keep a match, write the `score` to `1`. If you don't want the match, leave the `score` as anything but `1`

```{r}
matched %>%
  filter(score != 1) %>%
  arrange(desc(score)) %>%
  select(name, sector, name_ald, sector_ald, score) %>%
  head()
```

The final step is to prioritize your matches. This ensures that the same loan isn't accidentally matched to multiple `ald` companies. 

Read the validated match file back into R and run: 

```{r}
match_result <- matched %>%
  prioritize()

match_result %>%
  select(name_direct_loantaker, loan_size_outstanding, sector, name_ald) %>%
  head()
```

## r2dii.analysis

The final step is to analyze your matched dataset, and compare the results to a climate scenario. 

You can apply scenario targets to your loanbook data by calling: 

```{r}
library(r2dii.analysis)

portfolio_targets <- match_result %>%
  target_market_share(
    ald_demo,
    scenario_demo_2020,
    region_isos_demo
  )

portfolio_targets %>%
  head()
```

and plot the distribution of technologies using `ggplot2`:

```{r technology-mix}
library(ggplot2)

portfolio_targets %>%
  filter(
    sector == "power",
    weighted_production_metric != "normalized_corporate_economy",
    year == max(year)
  ) %>% # different targets can be plotted by setting these filters
  group_by(technology) %>%
  ggplot(aes(
    x = weighted_production_metric, y = weighted_production_value, fill =
      technology
  )) +
  geom_col(position = "fill") +
  labs(
    x = "Metric",
    y = "Weighted Capacity [%]"
  )
```
